---
title: "HW8"
author: "Zhuodiao Kuang"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(tidyverse)
library(knitr)
library(kableExtra)
library(survival)
library(flexsurv)
library(survminer)
library(survMisc)
```

# Problem 1

Since $T_i\sim LN(\mu,\sigma^2)$, we have the p.d.f. of the log-normal:
$$f(t) = \frac{1}{t\sigma\sqrt{2\pi}}\exp{\left(-\frac{(\ln t - \mu)}{2\sigma^2}\right)}$$
The log-likelihood of the sample $\{t_1,t_2,\ldots,t_n\}$ is:

$$LL = -\frac{n}{2}\ln(2\pi) - \frac{n}{2}\ln\sigma^2-\frac{1}{2\sigma^2}\sum_{i=1}^{n}(\ln t_i - \mu)^2 - \sum_{i=1}^{n}\ln t_i$$

Taking the F.O.C. with respect to $\mu$ and $\sigma$, and set them to zero, we can get:
$$
\begin{aligned}
\frac{\partial LL}{\partial \mu} & = \frac{1}{\sigma^2}\left(\sum_{i=1}^{n}\ln t_i - \mu\right) = 0\\
\frac{\partial LL}{\partial \sigma^2} & = -\frac{n}{2\sigma^2}+\frac{\sum_{i=1}^{n}(\ln t_i - \mu)^2 }{2\sigma^4} = 0
\end{aligned}
$$

We can solve them MLE $\hat\mu, \hat\sigma^2$ from the above equations:
$$
\begin{aligned}
\hat\mu &= \frac{1}{n}\sum_{i=1}^{n}\ln t_i\\
\hat\sigma^2 &= \frac{1}{n}\sum_{i=1}^{n}(\ln t_i - \hat\mu)^2
\end{aligned}
$$
Therefore, MLEs have a closed form.

# Problem 2

```{r}
ovarianfit_exp <- survreg(Surv(futime, fustat) ~ ecog.ps, data = ovarian, dist = "exponential")
summary(ovarianfit_exp)

ovarianfit_weib <- survreg(Surv(futime, fustat) ~ ecog.ps, data = ovarian, dist = "weibull")
summary(ovarianfit_weib)

```


Because both tests of two distributions generate relatively high p-values, so the survival data, by ecog status, doesn't follow the exponential distribution or the Weibull distribution.


Besides we can take $log{-logS(t)}$ and plot against log t.

```{r}
ovarian1<- ovarian |>
  filter(ecog.ps == 1)
ovarian2<- ovarian |>
  filter(ecog.ps == 2)

km.fit1 <- survfit(Surv(futime, fustat) ~ 1, data = ovarian1)
km_surv1 <- data.frame(time = km.fit1$time, est = km.fit1$surv)

km.fit2 <- survfit(Surv(futime, fustat) ~ 1, data = ovarian2)
km_surv2 <- data.frame(time = km.fit2$time, est = km.fit2$surv)

ggplot(km_surv1, aes(x = log (time), y = log(-log(est))))+
    geom_step() +
  labs(title = "LLg plot when ecog.ps = 1 ")
ggplot(km_surv2, aes(x = log(time), y = log(-log(est))))+
    geom_step()  +
  labs(title = "LLg plot when ecog.ps = 2 ")
```


Because the slop of a straight line is not 1, it may by considered as a Weibull distribution. But the size of data is limited and the plot is not so smooth, with the p-values calculated above, we are not sure whether the survival data follows the exponential distribution or the Weibull distribution.

# Problem 3

The likelihood of Weibull distribution is:
$$\begin{aligned} 
L(\beta) & =\prod_{i=1}^n h\left(T_i\mid Z_i\right)^{\Delta_i} S\left(T_i\right) \\ 
& = \prod_{i=1}^n \left(h_0(T_i)e^{\beta Z_i}\right)^{\Delta_i}e^{-\lambda T_i^\alpha}\\
& =\prod_{i=1}^n\left(\lambda \alpha T_i^{\alpha-1}e^{\beta Z_i}\right)^{\Delta_i} e^{-\lambda T_i^\alpha}\\
& = \left(\lambda \alpha \times16^{\alpha-1}e^{\beta}\right)^{1} e^{-\lambda \times 16^\alpha}\times  e^{-\lambda 
\times 20^\alpha} \\
&\times \left(\lambda \alpha \times12^{\alpha-1}\right)^{1} e^{-\lambda \times 12^\alpha} \times e^{-\lambda \times 14^\alpha}\\
& \times \left(\lambda \alpha \times11^{\alpha-1}\right)^{1} e^{-\lambda \times 11^\alpha}\\
& \times \left(\lambda \alpha \times9^{\alpha-1}e^{\beta}\right)^{1} e^{-\lambda \times 9^\alpha}\\
& = (\lambda\alpha)^4(16^{\alpha-1}12^{\alpha-1}11^{\alpha-1}9^{\alpha-1})e^{2\beta - \lambda(16^\alpha+20^\alpha+12^\alpha+14^\alpha+11^\alpha+9^\alpha)}\\
& = (\lambda\alpha)^4 19008^{\alpha-1} e^{2\beta - \lambda(16^\alpha+20^\alpha+12^\alpha+14^\alpha+11^\alpha+9^\alpha)}
\end{aligned}$$


# Problem 4

```{r aft_log_logistic}
leu_dat = readxl::read_excel("Datasets.xlsx", sheet = "Leukaemia")
leu_dat$trt = relevel(factor(leu_dat$trt), ref = "Control")
leuk.aft <- survreg(Surv(time, event == 1) ~ trt, leu_dat,
                      dist = "loglogistic")
summary(leuk.aft)
```

From the AFT model summary, we can see that the coefficient corresponds to the treatment effect is significant, which means that the treatment 6-MP has a significant effect on the survival of patients with Acute Myelogenous Leukemia.

The corresponding model is:
$$
\log T_i = 1.8927 + 1.2655\times I(\text{trt}_i=\text{ 6-MP})
$$
